---
title: "Using Seurat with BD Rhapsody multi-omic single-cell data"
output:
  html_document:
    df_print: paged
    #code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


```{r load-packages, warning=FALSE, results='hide'}
library(Seurat)
library(ggplot2)
library(dplyr)
```


## Load Rhapsody data in R

Load the RDS object create in the create_seurat_object_abseq.R script.

```{r}
norm.data <- readRDS('project_abseq_seurat.rds')
sct.data <- readRDS('project_abseq_seurat.rds')
```


## Quality control and selecting cells for further analysis

```{r eval = TRUE, warning=FALSE, fig.width = 12, fig.height = 6}
# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
# Note that other species references may use the pattern "^Mt. or "^mt."
norm.data[["percent.mt"]] <- PercentageFeatureSet(norm.data, pattern = "^MT.")

# Visualize QC metrics as a violin plot
VlnPlot(norm.data, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

```{r eval = TRUE, warning=FALSE}
norm.data <- subset(norm.data, subset = percent.mt < 20 & nFeature_RNA > 200 & nFeature_RNA < 4000)


norm.data = NormalizeData(norm.data)
norm.data <- FindVariableFeatures(norm.data, selection.method = "vst", nfeatures = 2000)

all.genes <- rownames(norm.data)
norm.data = ScaleData(norm.data, features = all.genes, model.use = "linear")

norm.data = RunPCA(norm.data, features = VariableFeatures(object=norm.data), verbose = F)

ElbowPlot(norm.data, ndims=40)

norm.data <- RunUMAP(norm.data, dims = 1:30, verbose = FALSE)

norm.data <- FindNeighbors(norm.data, dims = 1:30, verbose = FALSE)
norm.data <- FindClusters(norm.data, verbose = FALSE)
```

### Apply sctransform normalization

Note that this single command replaces NormalizeData(), ScaleData(), and FindVariableFeatures().
Transformed data will be available in the SCT assay, which is set as the default after running sctransform
During normalization, we can also remove confounding sources of variation, for example, mitochondrial mapping percentage

```{r}
sct.data[["percent.mt"]] <- PercentageFeatureSet(sct.data, pattern = "^MT.")

# run sctransform
sct.data <- SCTransform(sct.data, vars.to.regress = "percent.mt", verbose = FALSE)

```

The latest version of sctransform also supports using glmGamPoi package which substantially improves the speed of the learning procedure. It can be invoked by specifying method="glmGamPoi".

if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")

BiocManager::install("glmGamPoi")

sct.data <- SCTransform(sct.data, method = "glmGamPoi", vars.to.regress = "percent.mt", verbose = FALSE)


```{r}
# These are now standard steps in the Seurat workflow for visualization and clustering
sct.data <- RunPCA(sct.data, verbose = FALSE)
ElbowPlot(sct.data, ndims=40)

```
```{r}
sct.data <- RunUMAP(sct.data, dims = 1:30, verbose = FALSE)

sct.data <- FindNeighbors(sct.data, dims = 1:30, verbose = FALSE)
sct.data <- FindClusters(sct.data, verbose = FALSE)
```

Compare the normalized UMAP to the SCT UMAP.

```{r}
DimPlot(norm.data, label = TRUE)
DimPlot(sct.data, label = TRUE)

```

