---
title: "Using Seurat with BD Rhapsody multi-omic single-cell data"
output:
  html_document:
    df_print: paged
    #code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


```{r load-packages, warning=FALSE, results='hide'}
library(Seurat)
library(ggplot2)
library(dplyr)
```


## Load Rhapsody data in R

Load the RDS object create in the create_seurat_object_abseq.R script.

```{r}
combined.data <- readRDS('jasper_abseq_seurat.rds')
```


## Quality control and selecting cells for further analysis

This follows the [Guided Clustering Tutorial](https://satijalab.org/seurat/articles/pbmc3k_tutorial.html) for QC and normalization
and the [Weighted Nearest Neighbor Analysis](https://satijalab.org/seurat/articles/weighted_nearest_neighbor_analysis.html) for clustering.

Quality control steps are often performed in order to filter out low quality cells. In this example, we remove cells with:  

1. High mitochondrial counts
2. Abnormally high or low number of distinct expressed bioproducts (mRNA/AbSeq)

For Rhapsody data, high mitochondrial counts do not directly correlate to dying cells, and so the threshold for filtering
by mitochondrial counts may be higher than in the Seurat tutorial.

These filtering criteria are determined through the QC steps shown below:

```{r eval = TRUE, warning=FALSE, fig.width = 12, fig.height = 6}
# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
# Note that other species references may use the pattern "^Mt. or "^mt."
combined.data[["percent.mt"]] <- PercentageFeatureSet(combined.data, pattern = "^MT.")

# Visualize QC metrics as a violin plot
VlnPlot(combined.data, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

nFeature_RNA is the number of distinct bioproducts detected a cell  
nCount_RNA is the total number of molecules detected within a cell 

From the violin plot above, we can determine the filtering criteria of:

1. Remove cells with >20% mitochondria content;
2. Remove cells with <200 or >4000 bioproducts  (mRNA/AbSeq)


Use the command below to filter the Seurat Object:

```{r eval = TRUE, warning=FALSE}
combined.data <- subset(combined.data, subset = percent.mt < 20 & nFeature_RNA > 200 & nFeature_RNA < 4000)
```


## Normalizing the data

The Seurat NormalizeData default settings normalize the expression of each bioproduct by the total expression, multiply by a scale factor (10,000), and log transforms the result:

```{r eval = TRUE, warning=FALSE}
combined.data = NormalizeData(combined.data)
```

# Identification of highly variable features (feature selection)

Next, find a set of bioproducts that are highly variable between cells; the `nfeatures` parameter in the command indicates that it will return the 2000 most variable bioproducts.

```{r eval = TRUE, warning=FALSE, fig.width = 12, fig.height = 6}
combined.data <- FindVariableFeatures(combined.data)

```

## Scaling the data

Next, a linear transformation, or scaling, will be performed on the data. This step is required for the next PCA step:

```{r eval = TRUE, warning=FALSE, results='hide', message=FALSE}
all.genes <- rownames(combined.data)
combined.data = ScaleData(combined.data, features = all.genes, model.use = "linear")
```

## Perform linear dimensional reduction

Following the scaling, a linear dimensional reduction will be performed on the data; we will use the variable features calculated by the previous step here in the `features` parameter.

```{r eval = TRUE, warning=FALSE, results='hide', message=FALSE}
combined.data = RunPCA(combined.data, features = VariableFeatures(object=combined.data), verbose = F)

```

# Check dimensionality of the data

```{r}
ElbowPlot(combined.data, ndims=40)
```


Switch over to AbSeqs

```{r}
DefaultAssay(combined.data) <- 'ADT'

VariableFeatures(combined.data) <- rownames(combined.data[["ADT"]])
combined.data <- NormalizeData(combined.data, normalization.method = 'CLR', margin = 2)
combined.data <- ScaleData(combined.data)
combined.data <- RunPCA(combined.data, reduction.name='apca')

ElbowPlot(combined.data, ndims=40)
```

Here the workflow uses both the RNA and AbSeq data for clustering. 

```{r}
# Identify multimodal neighbors. These will be stored in the neighbors slot, 
# and can be accessed using combined.data[['weighted.nn']]
# The WNN graph can be accessed at combined.data[["wknn"]], 
# and the SNN graph used for clustering at combined.data[["wsnn"]]
# Cell-specific modality weights can be accessed at combined.data$RNA.weight
combined.data <- FindMultiModalNeighbors(
  combined.data, reduction.list = list("pca", "apca"), 
  dims.list = list(1:30, 1:18), modality.weight.name = "RNA.weight")

combined.data <- RunUMAP(combined.data, nn.name = "weighted.nn", reduction.name = "wnn.umap", reduction.key = "wnnUMAP_")
combined.data <- FindClusters(combined.data, graph.name = "wsnn", algorithm = 3, resolution = 2, verbose = FALSE)
```

UMAPs based on WNN.

```{r}
DimPlot(combined.data, reduction = 'wnn.umap', label = TRUE, repel = TRUE, label.size = 2.5) + ggtitle('Multimodal Clusters')
DimPlot(combined.data, reduction = 'wnn.umap', group.by = 'sample_name') + ggtitle('Multimodal Clusters')

```


Also take a look at the RNA and AbSeq UMAPs separately.

```{r}
combined.data <- RunUMAP(combined.data, reduction = 'pca', dims = 1:30, assay = 'RNA', 
              reduction.name = 'rna.umap', reduction.key = 'rnaUMAP_')
combined.data <- RunUMAP(combined.data, reduction = 'apca', dims = 1:18, assay = 'ADT', 
              reduction.name = 'adt.umap', reduction.key = 'adtUMAP_')

DimPlot(combined.data, reduction = 'rna.umap', group.by = 'sample_name') + ggtitle('RNA UMAP')
DimPlot(combined.data, reduction = 'adt.umap', group.by = 'sample_name') + ggtitle('AbSeq UMAP')

```


```{r}
FeaturePlot(combined.data, features = c("CD46.CD46.AHS0071.pAbO","CD34.8G12.CD34.AHS0182.pAbO"),
                  reduction = 'wnn.umap', max.cutoff = 2)

VlnPlot(combined.data, features = c("rna_CD3E","rna_CD8A"))

FeatureScatter(combined.data, feature1='CD34.8G12.CD34.AHS0182.pAbO', feature2='CD38.HIT2.CD38.AHS0022.pAbO', group.by = 'sample_name')

```
